---
title: "Reading hemibrainr data from google drive"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flywire}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> "We’re Building Something, Here, Detective, We’re Building It From Scratch. All The Pieces Matter."
([Lester, The Wire](https://en.wikipedia.org/wiki/The_Wire))

# Reading lots of data with hemibrainr

In order to use this package to its fullest, you need to get `hemibrainr` to read large amounts of data stored on a google drive. To do this, you have three options:

1. Mount your google drive usin [Google filestream](https://support.google.com/drive/answer/7329379?authuser=2)
2. Mount your google drive using [rclone](https://rclone.org/drive/)
3. Download the google drive and save locally

The best option is to use google filestream. By default, this is what `hemibrainr` expects. However, ou need a [Google Workspace](https://workspace.google.com/pricing.html) account (formerly G-Suite), which is a paid-for service. Without this, the best option is to use [rclone](https://rclone.org/drive/). You can also download to an external hard drive and use that.

We have two [Google team drives](https://support.google.com/a/users/answer/9310156?hl=en) available for you to use, which contain similar data. One ("hemibrain") is for internal use by the [Drosphila Connctomcis Group](https://www.zoo.cam.ac.uk/research/groups/connectomics). The other one ("hemibrainr") is shared with those who would like access. Contact us by email to request access.

# Option 1: Google workspace, hemibrainr and R

[Google drive](https://en.wikipedia.org/wiki/Google_Drive) is a data storage and synchonisation system. A Google team drive (shared drive) is a Google drive that is shared by a team of people, you can find them on your 'shared drives' in your Google authenticated account. Sharing and file ownership are managed for the entire team rather than one individual. Google Drive is a key component of [Google Workspace](https://workspace.google.com/pricing.html), Google's monthly subscription offering for businesses and organizations. As part of select Google Workspace plans, Drive offers unlimited storage, advanced file audit reporting, enhanced administration controls, and greater collaboration tools for teams.
[Google filestream](https://support.google.com/drive/answer/7329379?authuser=2) allows users with access to a Google Workspace to mount their Google Drives like a hard-drive, to their local machines. It is this feature that the package `hemibrain` capitalises upon in order to store large amounts of fly connectomics related data, and enable users to quickly access it.

On the hemibrain [Google team drive](https://support.google.com/a/users/answer/9310156?hl=en) for the [Drosphila Connctomcis Group](https://www.zoo.cam.ac.uk/research/groups/connectomics). If you do not have access to this team drive and would like to use it, to make the most our of `hemibrainr`, please get in contact. You will need top to have access to the drive and have [Google filestream](https://support.google.com/a/answer/7491144?hl=en) mounted. Then you will be able to:

* Read thousands of pre-skeletonised flywire/hemibrain neurons from Google Drive
* Read flywire/hemibrain NBLASTs and NLBLASTs to hemibrain neurons
* Read flywire/hemibrain neurons that are pre-transformed into a variety of brainspaces
* Get processed hemibrain connectivity data, synapse positions
* Get meta data for all flywire and hemibrain neurons
* Get information on flywire/FAFB/hemibrain neuron-neuron matches between data sets and hemispheres.

You can also set uo your own Google team drive in which to deposit data. We can help you get started, so get in contact. A starting point for creating thge data are the .R scripts in /data-raw/ of the `hemibrainr` Github repository. You will need the file structure from `hemibrainr:::hemibrainr_gdrive_structure`.

## Get your drive location

Let us have a look at the default `hemibrainr` settings, letting it to know where to find google drive data:

```{r see.paths, eval = FALSE}
# Load package
library(hemibrainr)

# This is where hemibrainr expects to find data, if not on google drive 
# or when you use the argument local = TRUE
options()$hemibrain_data

# Else, it wants to see it on the mounted team drive, here
options()$Gdrive_hemibrain_data

# Get hemibrain skeletons
db <- hemibrain_neurons(local=FALSE) 
## Will not work without google filestream

# You can also specify certain google sheets
## Used for neuron-neuron matching
options()$hemibrainr_matching_gsheet
```

## Set your drive location

If you need to set `hemibrainr` to look at a new drive, use:

```{r new.drive, eval = FALSE}
# Set a new google drive, can be the team drive name or a path to the correct drive
hemibrainr_set_drive("hemibrainr")

# Now just get the name of your default team drive.
## This will be used to locate your team drive using the R package googledrive
hemibrainr_team_drive()
```

# Option 2: rclone, hemibrainr and R
